\documentclass{book}
%\usepackage{cite}
\usepackage{harvard}
\input{26Snipes_macros}
\usepackage{alltt}
\usepackage{pdfpages}
\usepackage{enumitem}
\usepackage{url}
\begin{document}
\bibliographystyle{agsm}
\title{A Practical Guide to Analyzing IDE Usage Data\vspace{-0ex}}

\chapter{A Practical Guide to Analyzing IDE Usage Data\vspace{-0ex}}
\author{
Will Snipes, Emerson Murphy-Hill,
Thomas Fritz, Mohsen Vakilian \\
Kostadin Damevski,
Anil R. Nair, David Shepherd
}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\begin{center}
\section*{Abstract}
\end{center}
Integrated Development Environments (IDEs) such as Eclipse and Visual Studio provide tools and capabilities to perform tasks such as navigating among classes and methods, continuous compilation, code refactoring, automated testing, and integrated debugging all designed to increase productivity.  Instrumenting the IDE to collect usage data gives researchers a greater level of detail on developers' work than previously possible.  Usage data supports analysis of how developers spend their time, what activities might benefit from greater tool support, where developers have difficulty comprehending code, and whether they are following specific practices such as test-driven development.  With usage data, we expect to uncover more nuggets of how developers create mental models, how they investigate code, how they perform mini trial-and-error experiments, and what might drive productivity improvements for everyone.

\section{Introduction}
As software development evolved, many developers began using Integrated Development Environments (IDEs) to help manage the complexity of software programs.  Modern IDEs such as Eclipse and Visual Studio include tools and capabilities to improve developer productivity by assisting with tasks such as navigating among classes and methods, continuous compilation, code refactoring, automated testing, and integrated debugging.  The breadth of development activities supported by the IDE makes collecting editor, command, and tool usage data valuable for analyzing developers' work patterns.  

Instrumenting the IDE to collect usage data gives researchers a greater level of detail on developers' work than previously possible. Instrumenting the IDE involves extending the IDE within a provided API framework.  Eclipse and Visual Studio support a rich API framework allowing logging of many commands and actions as they occur.  Leveraging this API, researchers have built tools that observe all the commands developers use, developer actions within the editor such as browsing or inserting new code, and other add-in tools developers use.  Section~\ref{SecHowToCollectData} provides a how-to guide for researchers to use when implementing tools to collect data from Eclipse or Visual Studio.  

Collecting IDE usage data provides an additional view of how developers produce software that researchers can use to advance the practice of software engineering.  
The most obvious application of usage data is to analyze how developers spend their time in the IDE by classifying the events in the usage log and tracking time between each event.  Through usage data analysis, researchers and organizations gain a better understanding of the developer's time allocation and can identify opportunities to save time such as reducing build time or improving source code search and navigation tools.  Beyond time analysis, researchers have applied usage data to quantify developers use of practices such as the study of types of refactoring by~\citeasnoun{V:MurphyHill2012How} that found developers mostly perform minor refactoring while making other changes.  In another example,~\citeasnoun{Carter2010Are} leverage usage data to discover areas of the code where developers have difficulty with comprehension and should ask for assistance from more experienced developers.  One study determined whether developers are doing test driven development properly by writing the tests first then writing code to make those tests pass or are doing it improperly by writing tests against previously written code~\cite{Kou2010Operational}.  ~\citeasnoun{fritzBookChapter} describe how to collect and process data for recommendation systems including tools and analysis methods, and they discuss important findings from developer usage data analysis.    These works provide a good background and examples where usage data provided necessary information to answer interesting research questions in software engineering.

%For example, usage data supports analysis of how developers spend their time~\cite{V:johnson2003beyond}, developer actions that might benefit from greater tool support~\cite{V:MurphyHill2012How}, where developers have difficulty comprehending code~\cite{Carter2010Are}, even whether they are following specific practices such as test-driven development~\cite{Kou2010Operational}.   With these data, we can answer questions such as the level of expertise developers have for an area of source code~\cite{Fritz2010Degreeofknowledge}.   In another example, ~\citeasnoun{fritzBookChapter} describes how to collect and process data for recommendation systems including tools, analysis methods, and important findings from developer usage data analysis.    

There are limits, however, to what IDE usage data can tell us.  The missing elements include the developer's mental model of the code, and how they intend to alter the code to suit new requirements.  The developers' experience, design ideas, and constraints they keep in mind during an implementation activity are factors that we must obtain separately.  

Looking forward, usage data from development environments provides a platform for greater understanding of low-level developer practices.  We expect to uncover more nuggets of how developers work to comprehend source code, how they perform mini trial and error experiments, and what might release further productivity improvements for all developers.


\pagebreak


\input{26Snipes_UsageDataResearchConcepts}



Now that we have discussed aspects to consider when formulating your research plan, we are ready to dig deeper into specifics on how to collect data from developer IDEs.  The next Section covers several options that will propel your research forward.


\input{26Snipes_HowToCollectData}



Thus far we have been focusing on concrete usage data collection frameworks and the specific data collected by these frameworks.  With options to collect data from both Visual Studio and Eclipse, we hopefully provided a good resource to get you started towards your research goals.  Next, let's look next at methods and challenges in analyzing usage data.

\newpage

\input{26Snipes_HowToAnalyzeUsageData}


Using this overview of analysis techniques should give you a good start towards analyzing usage data for your research.
The ideas you bring to your usage data analysis process provide the real opportunities for innovating usage data based research.  Next we will discuss limitations we have observed in conducting research with usage data.

\section{Limits of What You Can Learn from Usage Data}
\label{sec:limitations}

Collecting usage data is potentially a boon to research, and as we have
pointed out, has many interesting and impactful applications.
Nonetheless, there are limits to what you can learn as a researcher
from usage data.
In fact, our experience is that researchers (and practitioners) have
high expectations about what they can learn from usage data, and those
expectations often come crashing down after significant effort implementing
and deploying a data collection system.
So before you begin your usage data collection and analysis, consider
the following two limitations of usage data.

\textbf{Rationale is Hard to Capture.}
Usage data tells you what a software developer did, but not
why she did it.
For example, if your usage data tells you that a developer used
new refactoring tool for the first time, from a trace alone you cannot determine whether
(a) she learned about the tool for the first time, (b) she had used the tool earlier, but before you started collecting data, or (c) her finger slipped and she pressed a hotkey by accident. We do not know whether she is satisfied with the the tool and will use it again in future.
A researcher may be able to distinguish between these by collecting additional information,
like asking the developer just after she used the tool why she used it,
but it is impossible to definitively separate these based on the
usage data alone.

\textbf{Usage Data Does Not Capture Everything.}
Researchers often want to capture ``everything,'' or at least
everything that a developer does.
This is practically impossible, and the reason we introduced the
goal-question metric.
If you have a system that captures all key presses, you are still
lacking information about mouse movements.
If you have a system that also captures mouse movements, you're still
missing the files that the developer is working with.
If your system captures also the files, you're still lacking the
histories of those files.
And so on.
In a theoretical sense, one could probably collect all observable information
about a programmer's behavior, yet the magnitude of effort required to 
do so would be enormous.
Furthermore, a significant amount of important information is probably not observable,
such as rationale and intent.
Ultimately, usage data is all about fitness for purpose -- is the data
you are analyzing fit for the purpose of your research questions?

To avoid these limitations, we recommend thinking systemically about how the data will be used while planning rather than thinking about usage data collection abstractly.
Invent an ideal usage data trace, and ask yourself:

\begin{itemize}[noitemsep]
  \item Does the data support my hypothesis?
  \item Are there alternative hypotheses that the data would support?
  \item Do I need additional data sources?
  \item Can the data be feasibly generated?
\end{itemize}

\noindent
Answering these questions will help you determine whether you can sidestep
the limitations of collecting and analyzing usage data.

\section{Conclusion}

Analyzing IDE usage data provides insights into many developer activities which help identify ways we can improve software engineering productivity.  For example, one study found developers spend more than half of their time browsing and reading source code (excluding editing and all other activities in the IDE)~\cite{SnipesExperiencesGamifyingSoftwareDevelopment}.  This finding supports initiatives to build and promote tools that improve the navigation experience by supporting structural navigation.  Other references mentioned in this chapter leveraged usage data to identify opportunities to improve refactoring tools~\cite{VakilianJohnson2014Alternate,MurphyHill2012Improving}.  By identifying task contexts through usage data, Mylyn made improvements to how developers manage their code and task context~\cite{Kersten-Mylyn}.  

If the last two decades could be labeled the era
of big data collection,
the next two decades will surely be labeled as the
era of smarter big data analysis.
Many questions still remain:
How do we balance data privacy and data richness?
What are the long term effects of developer monitoring?
How can we maximize the value of data collection
for as many researchers as possible, and reduce the
strain on research participants? How can we provide right data to right person at right time with the least effort?
Answering these questions will help our research
community advance in usage data collection and analysis.

Usage data, while now widely collected, still remains largely
an untapped resource by practitioners and researchers.
In this chapter, we have explained how to collect and
analyze usage data, which we hope will increase the ease
with which you can collect and analyze your own usage data.

% LocalWords: IDE IDEs UDC refactoring refactorings

\section{Acknowledgements}

This material is based in part upon work supported by the National 
Science Foundation under grant number 1252995.

\section{Code Listings}
\input{26Snipes_CodeListings}

\bibliographystyle{abbrv}
\bibliography{26Snipes_Bibliography} 


\end{document}
