\newcommand\CodingSpectator{\textsc{CodingSpectator}}

\newcommand\CodingTracker{\textsc{CodingTracker}}

\newcommand\Performed{\texttt{performed}}

\newcommand\Canceled{\texttt{canceled}}

\newcommand\Unavailable{\texttt{unavailable}}

\Def{NumberOfEclipseAutomatedRefactorings}{33}

\Def{NumberOfCodingSpectatorSupportedRefactorings}{23}

\subsection{\CodingSpectator}

\CodingSpectator~\cite{CodingSpectatorWebPage, VakilianETAL2011Richer,
VakilianETAL2012UseDisuseMisuse, VakilianETAL2013Compositional} is an extensible
framework for collecting Eclipse usage data. Although researchers at the
University of Illinois at Urbana-Champaign developed \CodingSpectator{}
primarily for collecting detailed data about the use of Eclipse refactoring
tool, it also provides a reusable infrastructure for \emph{submitting usage
data} from users to a central repository.
\CodingTracker~\cite{NegaraETAL2012Dangerous, NegaraETAL2013ManualRefactorings,
CodingTrackerWebPage} is another data collector developed at Illinois, which
reuses the data submission infrastructure provided by \CodingSpectator.

\subsubsection{What Data Is Collected}

\CodingSpectator{} was designed for capturing detailed data about the use of
automated refactorings. It collects three kinds of refactoring events:
\Canceled, \Performed, and \Unavailable. If a programmer starts an automated
refactoring but quits it before it finishes, \CodingSpectator{} records a
\Canceled{} refactoring event. If a programmer applies an automated refactoring,
\CodingSpectator{} records a \Performed{} refactoring event. Finally, if
programmer invokes an automated refactoring but the IDE refuses to start the
automated refactoring saying that the refactoring is not applicable to the
selected program element, \CodingSpectator{} records an \Unavailable{}
refactoring event.

Eclipse creates a \emph{refactoring descriptor} object for each \Performed{}
refactoring events and serializes it in an XML file. \CodingSpectator{} saves
more data in Eclipse refactoring descriptors of \Performed{} refactorings. In
addition, it creates and serializes refactoring descriptors for \Canceled{} and
\Unavailable{} refactoring events. \CodingSpectator{} supports
\Use{NumberOfCodingSpectatorSupportedRefactorings} of the
\Use{NumberOfEclipseAutomatedRefactorings} automated refactorings that Eclipse
supports.

\CodingSpectator{} records the following information in its refactoring
descriptors:

\begin{enumerate}

\item a timestamp recording when a refactoring event occurred

\item the location and contents of the selection that the programmer made before
  invoking the automated refactoring

\item the automated refactoring's identifier

\item configuration options, \eg{} input elements, project, and settings that
  programmers can set to control the effect of the refactoring

\item any problems reported by the automated refactoring to the programmer

\item the time that the programmer spent on each page of a refactoring wizard

\item whether the automated refactoring was invoked using Quick Assist

\end{enumerate}

% LocalWords: CodingSpectator Urbana Champaign IDE timestamp
%
% LocalWords: CodingTracker refactoring refactorings refactoring's

