\section{How to Collect Data}
\label{SecHowToCollectData}

There are many options for collecting usage data from IDEs.   Existing tools can provide ready made solutions for commonly used IDEs and some support collecting data from additional sources.   If you want to study data collected in prevous projects you can use the Eclipse archive for UDC data available below.

\subsection{Existing Data}

Existing UDC data is available through the Eclipse foundation at this URL:
\url{http://archive.eclipse.org/projects/usagedata/}
%could provide more analysis, maybe even BigData

\subsection{Instrumenting the IDE -- Overview of Frameworks} 

In this section we discuss details on how to use exisitng data collectors for Eclipse including Usage Data Collector, Mylyn Monitor, and Coding Spectator and provide a sample program to build your own data collector for Microsoft Visual Studio.  Before we get into details, here is an overview of some existing frameworks that may work for you.

\begin{itemize}

	\item Eclipse: Usage Data Collector (UDC) discussed in section \ref{EclipseUsageDataColector} collects commands executed in the environment and editors and views that are invoked.
	
	\item Eclipse: Mylyn Monitor described in section \ref{MylynMonitor} collects task oriented events and can be configured for saving data
	
	\item CodingSpectator %https://github.com/vazexqi/CodingSpectator
, discussed in section \ref{CodingSpectator} focuses on refactoring actions and the context in which they are taken.
	
	\item For Visual Studio, section \ref{buildItYourself} describes in detail how to build your own Visual Studio extension that collects all command events from the IDE.
	
	\item Hackystat provides a framework to collect usage data from many sources.  We do not discuss Hackystat here, hoever you can consult papers and the code website for it to collect more information. \cite{V:johnson2003beyond}, %https://code.google.com/p/hackystat/
	
	\item CodeAlike is a Visual Studio extension for personal analytics and research of usage data related to coding efficiency.  Currently CodeAlike is a proprietary tool whose developers seek to collaborate on usage data stuides.

\end{itemize}



\subsection{Eclipse Usage Data Collector}
\label{EclipseUsageDataCollector}
\input{EclipseUDC}

\subsection{Mylyn and the Eclipse Mylyn Monitor} 
\label{MylynMonitor}
\input{mylyn}

\input{codingspectator}

\subsection{Instrumenting the IDE -- Building it Yourself for Visual Studio} 
\label{buildItYourself}
\input{BuildItYourselfInVisualStudio}

%TODO Wrapup

